Endogenous Macrodynamics in Algorithmic Recourse

### 2022-08-09 dev/paper/paper.tex

\begin{document}
\maketitle
\begin{abstract}
Existing work on Counterfactual Explanations (CE) and Algorithmic
Recourse (AR) has largely been limited to the static setting: given some
classifier we are interested in finding close, actionable, realistic,
sparse, diverse and ideally causally founded counterfactuals. The
ability of CE to handle dynamics like data and model drift remains a
largely unexplored research challenge at this point. Only one recent
work considers the implications of exogenous domain and model shifts.
This project instead focuses on endogenous dynamics, that is shifts that
occur when AR is actually implemented by a proportion of individuals.
Early findings suggest that the involved shifts may be large with
important implications on the validity of AR and the overall
characteristics of the sample population.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, frame hidden, interior hidden, sharp corners]}{\end{tcolorbox}}\fi

@@

As mentioned earlier, research has so far also generally focused on
generating counterfactuals for single individuals or instances. We have
been able to identify only one existing work that investigates black-box
model behavior towards a group of individuals
\protect\hyperlink{ref-carrizosa2021generating}{\textbf{carrizosa2021generating?}}.
The authors propose an optimization framework that generates collective
counterfactuals. We provide a motivation for doing so from the
perspective of endogenous macrodynamics of algorithmic recourse.

@@

Much attention has been paid to the detection of dataset shifts --
situations where the distribution of data changes over time. Rabanser et
al.~suggest a framework to detect data drift from a minimal number of
samples through the application of two-sample tests
\protect\hyperlink{ref-rabanser2019failing}{{[}16{]}}. This task is a
generalization of the anomaly detection problem for large datasets,
which aims to answer the question if two sets of samples could have been
generated from the same probability distribution. Numerous approaches to
anomaly detection have been summarized
\protect\hyperlink{ref-chandola2009anomaly}{{[}17{]}}. Another
well-established research topic is that of concept drift -- situations
where external variables influence the patterns between the input and
the output of a model
\protect\hyperlink{ref-widmer1996learning}{{[}18{]}}. For instance, Gama
et al.~offer a review of the adaptive learning techniques which can
handle concept drift \protect\hyperlink{ref-gama2014survey}{{[}19{]}}.
Less previous work is available on the related topic of model drift -
changes in model performance over time. Nelson et al.~review how
resistant different machine learning models are to the model drift
\protect\hyperlink{ref-nelson2015evaluating}{{[}20{]}}. Ackerman et
al.~offer a method to detect changes in model performance when ground
truth is not available
\protect\hyperlink{ref-ackerman2021machine}{{[}21{]}}.

@@

Despite the large and growing number of approaches to counterfactual
search there have been surprisingly few benchmark studies that compare
different methodologies. This may be partially due to limited software
availability in this space. Recent work has started to address this gap:
firstly, \protect\hyperlink{ref-de2021framework}{{[}22{]}} run a large
benchmarking study using different algorithmic aproaches and numerous
tabular datasets; secondly,
\protect\hyperlink{ref-pawelczyk2021carla}{{[}4{]}} introduce a Python
framework that can be used to apply and benchmark different
methodologies; finally,
\protect\hyperlink{ref-altmeyer2022CounterfactualExplanations}{{[}6{]}}
provides an extensible, fast and language-agnostic implementation in
Julia. Since the experiments presented here involve extensive
simulations, we have relied on and extended the Julia implementation due
to the associated performance benefits.

@@

We use 6 synthetic binary classification datasets consisting of 200-400
samples grouped in normally-distributed clusters.\footnote{To see how
  the data is generated see here:
  \url{https://github.com/pat-alt/AlgorithmicRecourseDynamics.jl/blob/main/notebooks/synthetic_datasets.ipynb}}
The datasets are presented in Figure~\ref{fig-synthetic-data} (see also
Appendix A for a formal description). Samples from the negative class
are marked in blue while samples of the positive class are marked in
orange.

@@

{\centering \includegraphics{www/synthetic_data.png}

@@

Additionally, we use two real-world datasets from the Finance domain.
Firstly, we use the Give Me Some Credit dataset which was open-sourced
on Kaggle for the task to predict whether a borrower is likely to
experience financial difficulties in the next two years
\protect\hyperlink{ref-gmsc_data}{{[}23{]}}. Originally consisting of
250,000 instances with 11 numerical attributes, the dataset was randomly
undersampled to result in a balanced subsample made up of 3000
individuals. Secondly, we the German Credit dataset which involves the
task of predicting if bank customers are credit-worthy or not
\protect\hyperlink{ref-germancredit1994}{{[}24{]}}. It consists of 700
positive and 300 negative instances charaterized by 7 numerical and 13
categorical attributes. We process the dataset in two ways: (1) the
values of the ``Personal status and sex'' feature are aggregated by the
two represented genders; (2) the most common values are calculated for
all categorical features such that a feature \(x_d\) with the mode
\(\bar{x}_d\) is transformed into a new binary feature
\(\tilde{x}_d=\mathbb{I}_{x_{d,i}>=\bar{x}_d}\). Binarization ensures
that we can use all counterfactual generators in the benchmark.

@@

The evaluation metric in Equation~\ref{eq-mmd} is computed after every
round \(t=1,...,T\) of the experiment. To assess the statistical
significance of the observed shifts under the null hypothesis that
samples \(X\) and \(\tilde{X}\) were drawn from the same probability
distribution we follow
\protect\hyperlink{ref-arcones1992bootstrap}{{[}25{]}}. To that end, we
combine the two samples and generate a large number of permutations of
\(X + \tilde{X}\). Then, we split the permuted data into two new samples
\(X^\prime\) and \(\tilde{X}^\prime\) having the same size as the
original samples. Then under the null hypothesis we should have that
\(MMD_u^2[F,X^\prime,\tilde{X}^\prime]\) be approximately equal to
\(MMD_u^2[F,X,\tilde{X}]\). The corresponding \(p\)-value can then be
calculated by counting how these two quantities are not equal.

@@

As an alternative to PP MMD we use a pseudo-distance for the
\textbf{Disagreement Coefficient} (Disagreement). This metric was
introduced in \protect\hyperlink{ref-hanneke2007bound}{{[}26{]}} and
estimates \(p(M(x) \neq M^\prime(x))\), that is the probability that two
classifiers do not agree on the predicted outcome for a randomly chosen
sample. Thus, it is not relevant whether the classification is correct
according to the ground truth, but only whether the sample lies on the
same side of the two respective decision boundaries. In our context,
this metric quantifies the overlap between the initial model (trained
before the application of recourse) and the updated model. A
Disagreement Coefficient unequal to zero is indicative of a model shift.
The opposite is not true: even if the Disagreement Coefficient is equal
to zero a model shift may still have occured. This is one reason for why
PP MMD is our our preferred metric.

@@

For reasons stated earlier we have limited our analysis to
differentiable linear and non-linear classifiers, in particular logistic
regression and deep neural networks. While these sorts of classifiers
have also typically been analyzed in the existing literature on
counterfactual explanations and algorithmic recourse, they represent
only a subset of popular machine learning models employed in practice -
both black-box and glass-box. Despite the success and popularity of deep
learning in the context of high-dimensional data such as image, audio
and video, empirical evidence suggests that other models such as boosted
decision trees may have an edge when it comes to lower-dimensional
tabular datasets, such as the ones considered here
\protect\hyperlink{ref-grinsztajn2022tree}{{[}28{]}}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-rabanser2019failing}{}}%
\CSLLeftMargin{{[}16{]} }%
\CSLRightInline{S. Rabanser, S. Günnemann, and Z. Lipton, {``Failing
loudly: An empirical study of methods for detecting dataset shift,''}
\emph{Advances in Neural Information Processing Systems}, vol. 32,
2019.}

\leavevmode\vadjust pre{\hypertarget{ref-chandola2009anomaly}{}}%
\CSLLeftMargin{{[}17{]} }%
\CSLRightInline{V. Chandola, A. Banerjee, and V. Kumar, {``Anomaly
detection: A survey,''} \emph{ACM computing surveys (CSUR)}, vol. 41,
no. 3, pp. 1--58, 2009.}

\leavevmode\vadjust pre{\hypertarget{ref-widmer1996learning}{}}%
\CSLLeftMargin{{[}18{]} }%
\CSLRightInline{G. Widmer and M. Kubat, {``Learning in the presence of
concept drift and hidden contexts,''} \emph{Machine learning}, vol. 23,
no. 1, pp. 69--101, 1996.}

\leavevmode\vadjust pre{\hypertarget{ref-gama2014survey}{}}%
\CSLLeftMargin{{[}19{]} }%
\CSLRightInline{J. Gama, I. Žliobaitė, A. Bifet, M. Pechenizkiy, and A.
Bouchachia, {``A survey on concept drift adaptation,''} \emph{ACM
computing surveys (CSUR)}, vol. 46, no. 4, pp. 1--37, 2014.}

\leavevmode\vadjust pre{\hypertarget{ref-nelson2015evaluating}{}}%
\CSLLeftMargin{{[}20{]} }%
\CSLRightInline{K. Nelson, G. Corbin, M. Anania, M. Kovacs, J. Tobias,
and M. Blowers, {``Evaluating model drift in machine learning
algorithms,''} in \emph{2015 IEEE symposium on computational
intelligence for security and defense applications (CISDA)}, 2015, pp.
1--8.}

\leavevmode\vadjust pre{\hypertarget{ref-ackerman2021machine}{}}%
\CSLLeftMargin{{[}21{]} }%
\CSLRightInline{S. Ackerman, P. Dube, E. Farchi, O. Raz, and M.
Zalmanovici, {``Machine learning model drift detection via weak data
slices,''} in \emph{2021 IEEE/ACM third international workshop on deep
learning for testing and testing for deep learning (DeepTest)}, 2021,
pp. 1--8.}

\leavevmode\vadjust pre{\hypertarget{ref-de2021framework}{}}%
\CSLLeftMargin{{[}22{]} }%
\CSLRightInline{R. M. B. de Oliveira and D. Martens, {``A framework and
benchmarking study for counterfactual generating methods on tabular
data,''} \emph{Applied Sciences}, vol. 11, no. 16, p. 7274, 2021.}

\leavevmode\vadjust pre{\hypertarget{ref-gmsc_data}{}}%
\CSLLeftMargin{{[}23{]} }%
\CSLRightInline{K. Competition, {``Give me some credit, improve on the
state of the art in credit scoring by predicting the probability that
somebody will experience financial distress in the next two years.''}
\url{https://www.kaggle.com/c/GiveMeSomeCredit}}

\leavevmode\vadjust pre{\hypertarget{ref-germancredit1994}{}}%
\CSLLeftMargin{{[}24{]} }%
\CSLRightInline{H. Hoffman, {``German credit data,''} 1994.
\url{https://archive.ics.uci.edu/ml/datasets/statlog+(german+credit+data)}}

\leavevmode\vadjust pre{\hypertarget{ref-arcones1992bootstrap}{}}%
\CSLLeftMargin{{[}25{]} }%
\CSLRightInline{M. A. Arcones and E. Gine, {``On the bootstrap of u and
v statistics,''} \emph{The Annals of Statistics}, pp. 655--674, 1992.}

\leavevmode\vadjust pre{\hypertarget{ref-hanneke2007bound}{}}%
\CSLLeftMargin{{[}26{]} }%
\CSLRightInline{S. Hanneke, {``A bound on the label complexity of
agnostic active learning,''} in \emph{Proceedings of the 24th
international conference on machine learning}, 2007, pp. 353--360.}

\leavevmode\vadjust pre{\hypertarget{ref-borisov2021deep}{}}%
\CSLLeftMargin{{[}27{]} }%
\CSLRightInline{V. Borisov, T. Leemann, K. Seßler, J. Haug, M.
Pawelczyk, and G. Kasneci, {``Deep neural networks and tabular data: A
survey,''} \emph{arXiv preprint arXiv:2110.01889}, 2021.}

\leavevmode\vadjust pre{\hypertarget{ref-grinsztajn2022tree}{}}%
\CSLLeftMargin{{[}28{]} }%
\CSLRightInline{L. Grinsztajn, E. Oyallon, and G. Varoquaux, {``Why do
tree-based models still outperform deep learning on tabular data?''}
\emph{arXiv preprint arXiv:2207.08815}, 2022.}

>>>>

\begin{document}
\maketitle
\begin{abstract}
Existing work on Counterfactual Explanations (CE) and Algorithmic
Recourse (AR) has largely been limited to the static setting: given some
classifier we are interested in finding close, actionable, realistic,
sparse, diverse and ideally causally founded counterfactuals. The
ability of CE to handle dynamics like data and model drift remains a
largely unexplored research challenge at this point. Only one recent
work considers the implications of exogenous domain and model shifts.
This project instead focuses on endogenous dynamics, that is shifts that
occur when AR is actually implemented by a proportion of individuals.
Early findings suggest that the involved shifts may be large with
important implications on the validity of AR and the overall
characteristics of the sample population.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, breakable, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi

@@

As mentioned earlier, research has so far also generally focused on
generating counterfactuals for single individuals or instances. We have
been able to identify only one existing work that investigates black-box
model behavior towards a group of individuals
\protect\hyperlink{ref-carrizosa2021generating}{{[}16{]}}. The authors
propose an optimization framework that generates collective
counterfactuals. We provide a motivation for doing so from the
perspective of endogenous macrodynamics of algorithmic recourse.

@@

Much attention has been paid to the detection of dataset shifts --
situations where the distribution of data changes over time. Rabanser et
al.~suggest a framework to detect data drift from a minimal number of
samples through the application of two-sample tests
\protect\hyperlink{ref-rabanser2019failing}{{[}17{]}}. This task is a
generalization of the anomaly detection problem for large datasets,
which aims to answer the question if two sets of samples could have been
generated from the same probability distribution. Numerous approaches to
anomaly detection have been summarized
\protect\hyperlink{ref-chandola2009anomaly}{{[}18{]}}. Another
well-established research topic is that of concept drift -- situations
where external variables influence the patterns between the input and
the output of a model
\protect\hyperlink{ref-widmer1996learning}{{[}19{]}}. For instance, Gama
et al.~offer a review of the adaptive learning techniques which can
handle concept drift \protect\hyperlink{ref-gama2014survey}{{[}20{]}}.
Less previous work is available on the related topic of model drift -
changes in model performance over time. Nelson et al.~review how
resistant different machine learning models are to the model drift
\protect\hyperlink{ref-nelson2015evaluating}{{[}21{]}}. Ackerman et
al.~offer a method to detect changes in model performance when ground
truth is not available
\protect\hyperlink{ref-ackerman2021machine}{{[}22{]}}.

@@

Despite the large and growing number of approaches to counterfactual
search there have been surprisingly few benchmark studies that compare
different methodologies. This may be partially due to limited software
availability in this space. Recent work has started to address this gap:
firstly, \protect\hyperlink{ref-de2021framework}{{[}23{]}} run a large
benchmarking study using different algorithmic aproaches and numerous
tabular datasets; secondly,
\protect\hyperlink{ref-pawelczyk2021carla}{{[}4{]}} introduce a Python
framework that can be used to apply and benchmark different
methodologies; finally,
\protect\hyperlink{ref-altmeyer2022CounterfactualExplanations}{{[}6{]}}
provides an extensible, fast and language-agnostic implementation in
Julia. Since the experiments presented here involve extensive
simulations, we have relied on and extended the Julia implementation due
to the associated performance benefits.

@@

We use four synthetic binary classification datasets consisting of 1000
samples each.\footnote{To see how the data is generated see here:
  \url{https://github.com/pat-alt/AlgorithmicRecourseDynamics.jl/blob/main/notebooks/synthetic_datasets.ipynb}}
The datasets are presented in Figure~\ref{fig-synthetic-data} (see also
Appendix A for a formal description). Samples from the negative class
are marked in blue while samples of the positive class are marked in
orange.

@@

{\centering \includegraphics[width=8cm,height=8cm]{www/placeholder.png}

@@

Additionally, we use two real-world datasets from the Finance domain.
Firstly, we use the Give Me Some Credit dataset which was open-sourced
on Kaggle for the task to predict whether a borrower is likely to
experience financial difficulties in the next two years
\protect\hyperlink{ref-gmsc_data}{{[}24{]}}. Originally consisting of
250,000 instances with 11 numerical attributes, the dataset was randomly
undersampled to result in a balanced subsample made up of 3000
individuals. Secondly, we the German Credit dataset which involves the
task of predicting if bank customers are credit-worthy or not
\protect\hyperlink{ref-germancredit1994}{{[}25{]}}. It consists of 700
positive and 300 negative instances charaterized by 7 numerical and 13
categorical attributes. We process the dataset in two ways: (1) the
values of the ``Personal status and sex'' feature are aggregated by the
two represented genders; (2) the most common values are calculated for
all categorical features such that a feature \(x_d\) with the mode
\(\bar{x}_d\) is transformed into a new binary feature
\(\tilde{x}_d=\mathbb{I}_{x_{d,i}>=\bar{x}_d}\). Binarization ensures
that we can use all counterfactual generators in the benchmark.

@@

The evaluation metric in Equation~\ref{eq-mmd} is computed after every
round \(t=1,...,T\) of the experiment. To assess the statistical
significance of the observed shifts under the null hypothesis that
samples \(X\) and \(\tilde{X}\) were drawn from the same probability
distribution we follow
\protect\hyperlink{ref-arcones1992bootstrap}{{[}26{]}}. To that end, we
combine the two samples and generate a large number of permutations of
\(X + \tilde{X}\). Then, we split the permuted data into two new samples
\(X^\prime\) and \(\tilde{X}^\prime\) having the same size as the
original samples. Then under the null hypothesis we should have that
\(MMD_u^2[F,X^\prime,\tilde{X}^\prime]\) be approximately equal to
\(MMD_u^2[F,X,\tilde{X}]\). The corresponding \(p\)-value can then be
calculated by counting how these two quantities are not equal.

@@

As an alternative to PP MMD we use a pseudo-distance for the
\textbf{Disagreement Coefficient} (Disagreement). This metric was
introduced in \protect\hyperlink{ref-hanneke2007bound}{{[}27{]}} and
estimates \(p(M(x) \neq M^\prime(x))\), that is the probability that two
classifiers do not agree on the predicted outcome for a randomly chosen
sample. Thus, it is not relevant whether the classification is correct
according to the ground truth, but only whether the sample lies on the
same side of the two respective decision boundaries. In our context,
this metric quantifies the overlap between the initial model (trained
before the application of recourse) and the updated model. A
Disagreement Coefficient unequal to zero is indicative of a model shift.
The opposite is not true: even if the Disagreement Coefficient is equal
to zero a model shift may still have occured. This is one reason for why
PP MMD is our our preferred metric.

@@

For reasons stated earlier we have limited our analysis to
differentiable linear and non-linear classifiers, in particular logistic
regression and deep neural networks. While these sorts of classifiers
have also typically been analyzed in the existing literature on
counterfactual explanations and algorithmic recourse, they represent
only a subset of popular machine learning models employed in practice -
both black-box and glass-box. Despite the success and popularity of deep
learning in the context of high-dimensional data such as image, audio
and video, empirical evidence suggests that other models such as boosted
decision trees may have an edge when it comes to lower-dimensional
tabular datasets, such as the ones considered here
\protect\hyperlink{ref-grinsztajn2022tree}{{[}29{]}}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-carrizosa2021generating}{}}%
\CSLLeftMargin{{[}16{]} }%
\CSLRightInline{E. Carrizosa, J. Ramırez-Ayerbe, and D. Romero,
{``Generating collective counterfactual explanations in score-based
classification via mathematical optimization,''} 2021.}

\leavevmode\vadjust pre{\hypertarget{ref-rabanser2019failing}{}}%
\CSLLeftMargin{{[}17{]} }%
\CSLRightInline{S. Rabanser, S. Günnemann, and Z. Lipton, {``Failing
loudly: An empirical study of methods for detecting dataset shift,''}
\emph{Advances in Neural Information Processing Systems}, vol. 32,
2019.}

\leavevmode\vadjust pre{\hypertarget{ref-chandola2009anomaly}{}}%
\CSLLeftMargin{{[}18{]} }%
\CSLRightInline{V. Chandola, A. Banerjee, and V. Kumar, {``Anomaly
detection: A survey,''} \emph{ACM computing surveys (CSUR)}, vol. 41,
no. 3, pp. 1--58, 2009.}

\leavevmode\vadjust pre{\hypertarget{ref-widmer1996learning}{}}%
\CSLLeftMargin{{[}19{]} }%
\CSLRightInline{G. Widmer and M. Kubat, {``Learning in the presence of
concept drift and hidden contexts,''} \emph{Machine learning}, vol. 23,
no. 1, pp. 69--101, 1996.}

\leavevmode\vadjust pre{\hypertarget{ref-gama2014survey}{}}%
\CSLLeftMargin{{[}20{]} }%
\CSLRightInline{J. Gama, I. Žliobaitė, A. Bifet, M. Pechenizkiy, and A.
Bouchachia, {``A survey on concept drift adaptation,''} \emph{ACM
computing surveys (CSUR)}, vol. 46, no. 4, pp. 1--37, 2014.}

\leavevmode\vadjust pre{\hypertarget{ref-nelson2015evaluating}{}}%
\CSLLeftMargin{{[}21{]} }%
\CSLRightInline{K. Nelson, G. Corbin, M. Anania, M. Kovacs, J. Tobias,
and M. Blowers, {``Evaluating model drift in machine learning
algorithms,''} in \emph{2015 IEEE symposium on computational
intelligence for security and defense applications (CISDA)}, 2015, pp.
1--8.}

\leavevmode\vadjust pre{\hypertarget{ref-ackerman2021machine}{}}%
\CSLLeftMargin{{[}22{]} }%
\CSLRightInline{S. Ackerman, P. Dube, E. Farchi, O. Raz, and M.
Zalmanovici, {``Machine learning model drift detection via weak data
slices,''} in \emph{2021 IEEE/ACM third international workshop on deep
learning for testing and testing for deep learning (DeepTest)}, 2021,
pp. 1--8.}

\leavevmode\vadjust pre{\hypertarget{ref-de2021framework}{}}%
\CSLLeftMargin{{[}23{]} }%
\CSLRightInline{R. M. B. de Oliveira and D. Martens, {``A framework and
benchmarking study for counterfactual generating methods on tabular
data,''} \emph{Applied Sciences}, vol. 11, no. 16, p. 7274, 2021.}

\leavevmode\vadjust pre{\hypertarget{ref-gmsc_data}{}}%
\CSLLeftMargin{{[}24{]} }%
\CSLRightInline{K. Competition, {``Give me some credit, improve on the
state of the art in credit scoring by predicting the probability that
somebody will experience financial distress in the next two years.''}
\url{https://www.kaggle.com/c/GiveMeSomeCredit}}

\leavevmode\vadjust pre{\hypertarget{ref-germancredit1994}{}}%
\CSLLeftMargin{{[}25{]} }%
\CSLRightInline{H. Hoffman, {``German credit data,''} 1994.
\url{https://archive.ics.uci.edu/ml/datasets/statlog+(german+credit+data)}}

\leavevmode\vadjust pre{\hypertarget{ref-arcones1992bootstrap}{}}%
\CSLLeftMargin{{[}26{]} }%
\CSLRightInline{M. A. Arcones and E. Gine, {``On the bootstrap of u and
v statistics,''} \emph{The Annals of Statistics}, pp. 655--674, 1992.}

\leavevmode\vadjust pre{\hypertarget{ref-hanneke2007bound}{}}%
\CSLLeftMargin{{[}27{]} }%
\CSLRightInline{S. Hanneke, {``A bound on the label complexity of
agnostic active learning,''} in \emph{Proceedings of the 24th
international conference on machine learning}, 2007, pp. 353--360.}

\leavevmode\vadjust pre{\hypertarget{ref-borisov2021deep}{}}%
\CSLLeftMargin{{[}28{]} }%
\CSLRightInline{V. Borisov, T. Leemann, K. Seßler, J. Haug, M.
Pawelczyk, and G. Kasneci, {``Deep neural networks and tabular data: A
survey,''} \emph{arXiv preprint arXiv:2110.01889}, 2021.}

\leavevmode\vadjust pre{\hypertarget{ref-grinsztajn2022tree}{}}%
\CSLLeftMargin{{[}29{]} }%
\CSLRightInline{L. Grinsztajn, E. Oyallon, and G. Varoquaux, {``Why do
tree-based models still outperform deep learning on tabular data?''}
\emph{arXiv preprint arXiv:2207.08815}, 2022.}

### 2022-08-09 dev/paper/sections/methodology.qmd

We use 6 synthetic binary classification datasets consisting of 200-400 samples grouped in normally-distributed clusters.^[To see how the data is generated see here: [https://github.com/pat-alt/AlgorithmicRecourseDynamics.jl/blob/main/notebooks/synthetic_datasets.ipynb](https://github.com/pat-alt/AlgorithmicRecourseDynamics.jl/blob/main/notebooks/synthetic_datasets.ipynb)] The datasets are presented in @fig-synthetic-data (see also Appendix A for a formal description). Samples from the negative class are marked in blue while samples of the positive class are marked in orange.

![PLACEHOLDER: A visualization of the synthetic classification datasets used in our experiments.](www/synthetic_data.png){#fig-synthetic-data}

>>>>

We use four synthetic binary classification datasets consisting of 1000 samples each.^[To see how the data is generated see here: [https://github.com/pat-alt/AlgorithmicRecourseDynamics.jl/blob/main/notebooks/synthetic_datasets.ipynb](https://github.com/pat-alt/AlgorithmicRecourseDynamics.jl/blob/main/notebooks/synthetic_datasets.ipynb)] The datasets are presented in @fig-synthetic-data (see also Appendix A for a formal description). Samples from the negative class are marked in blue while samples of the positive class are marked in orange.

![PLACEHOLDER: A visualization of the synthetic classification datasets used in our experiments.](www/placeholder.png){#fig-synthetic-data fig.pos="h" width="8cm" height="8cm"}
